---
title: 'Lab 5: Between Subjects ANOVA'
author: "INSERT YOUR NAME HERE"
date: "2023-02-12"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(psych)
library(plyr) # install this package if you need to using install.packages(plyr)
library(dplyr)
```

Please be sure to get into pairs or trios and help each other. 

Purpose: 

The purpose of this lab is to review how to use R Studio to enter and analyze a one-way between-subjects ANOVA. 

Instructions: 

Prof. Motivation collected data from 12 participants (N=12) in a study comparing conceptual instruction to procedural instruction. In the study, participants solved a set of 60 mathematics problems after receiving one of two lessons. A third group did not receive a lesson, so this is our control group.

The conceptual lesson focused on a mathematical concept important for understanding the problems. The procedural lesson focused on a specific problem-solving procedure that could be used to solve the problems. In the control condition, participants did not receive a lesson before solving the problems. The dataset shows the number of problems solved correctly (out of 60) by each participant.

1. Download the dataset: “week5_dataset.csv”. 



2. Put the dataset into the folder you want to use for class today. This will be your working directory. 



3. Open R studio. Set your working directory. There are two ways to do this.

First way: Session -> Set Working Directory -> Choose Directory 

OR 

Enter the following code into the lower console of R studio. 
setwd(“INSERT FILEPATH HERE”) 



```{r Import data}
# import the dataset 
dat1 <- read.csv("week5_dataset.csv") # imports the data from the working directory
```

```{r View data}
# This is a small dataset so let's view the entire dataframe
# It should open in a new window 
View(dat1)
```

4. What is the null hypothesis? State it here: 





```{r Descriptive Statistics}
# This is a very fancy way of getting our descriptive statistics based on the condition. It uses the 'dplyr' package and calculates the mean, standard deviation and variance of each condition. 
group_by(dat1, Condition) %>%
  summarise(
    count = n(),
    mean = mean(Score),
    sd = sd(Score),
    var = var(Score)
  )
```


5. What is the mean of each condition (Conceptual, Procedural, Control)? What can you conclude? 




6. What is the variance of each condition? What does the variance of each condition tell you?



```{r One-Way ANOVA}
# Calculates a one-way between subject ANOVA
one.way <- aov(Score ~ Condition, data = dat1)
summary(one.way)
```



7. What is the p-value for the effect of condition?  What does that p-value mean?  Can we use these results to conclude that conceptual instruction is better than procedural instruction? Why or why not? (Hint: Do you know the difference between each group?)



8. What would you have to do to be able to make a conclusion about the difference between conceptual and procedural instruction? 



```{r Multiple Pairwise-Comparison}
# diff: difference between means of the two groups
# lwr, upr: the lower and the upper end point of the confidence interval at 95% (default)
# p adj: p-value after adjustment for the multiple comparisons.
TukeyHSD(one.way)
```

9. What conclusions can you state about the differences between each condition? 



10. Click on the arrow next to "Knit" and save either as a PDF, Word or HTML. Name the document "yourname_lab5".




